Constraining Black Hole Growth with High-z Quasar Observations

### 2022-09-23 ms.tex
.
version0923; from me; revised all of Masa's 2nd round comments (w/o proximity zone); next: integrate w/ k's revision

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
The early evolution of the quasar luminosity function (QLF) and black hole mass function (BHMF) encodes 
key information on the physics determining the radiative and accretion processes of supermassive 
black holes (SMBHs) in high-redshift quasars.
Although the shape of the QLF has been constrained by recent wide and deep surveys,
it still remains challenging to develop a theoretical model that explains its redshift evolution with BH growth self-consistently.
In this study, based on a semi-analytical model for the formation and growth of seed BHs
in quasar progenitor halos,
we construct the QLF and BHMF of the early BH population that experiences multiple accretion bursts,
in each of which a constant Eddington ratio is assigned following a Schechter distribution function.
Our best-fit model to reproduce the observed QLF and BHMF at $z\simeq 6$
suggests that several episodes of modest super-Eddington accretion occur
and each of them lasts $\tau \simeq 20-30$ Myr.
The average duty cycle of BHs that reach $\gtrsim 10^8~\Msun$ by $z\simeq 6$ is as high as $\simeq 15~\%$,
which is nearly twice that of the entire population.
We also find that the observed Eddington-ratio distribution function is skewed to a log-normal shape from the intrinsic Shechter-like function
owing to detection limits of quasar surveys.
The predicted redshift evolution of both QLF and BHMF of the earliest BHs suggests a rapid decay of their number and mass
toward higher redshifts of $z>6$, compared to galaxy stellar mass.
This will be unveiled by future deep and wide surveys with the James Webb Space Telescope, Roman Space Telescope, and Euclid.
\end{abstract}

@@

\orange{
We consider seed BH formation as PopIII stellar remnants in the main progenitors of quasar host galaxies.
We note that the more abundant but less massive PopII stellar remnants are likely to dominate
the low mass end of the BHMF at $z\sim 6$,
thus not significantly contribute to the observable QLF and BHMF \citep{2022MNRAS.511..616T}.
}
The BH seeding time and mass are determined adpoting a semi-analytical model established by \citet{2021ApJ...917...60L}.
In the highly-biased, overdense regions of the universe, those progenitor halos are likely irradiated by intense H$_2$-photodissociating radiation 
from nearby star-forming galaxies and heat the interior gas by successive mergers. 
The two effects counteracting H$_2$ formation and cooling prevent gas collapse and delay prior star formation \citep[e.g.,][]{2014MNRAS.445..107V,2019Natur.566...85W}.
Under such peculiar circumstances, massive clouds collapse to the halo centers at high accretion rates and leave massive stars (presumably seed BHs) behind.
Our model takes into account various key physical processes on star formation such as halo merger heating, radiative cooling, and
(photo-)chemical reaction networks.
%, and gravitational instability of the central gaseous cores.
The time evolution of the gas dynamics and thermal states is calculated in a self-consistent way with the halo assembly history.
%the gas dynamical and thermal evolution is regulated by the parent halo growth histories, and the environmental streaming velocities.
%
%Next we introduce our seeding model where the gas dynamical and thermal evolution is regulated by the parent halo growth histories, 
%and the environmental streaming velocities.
%In \citet{2021ApJ...917...60L}, we propose an semi-analytical model which 
%traces the gas evolution along its parent halo growth and yields the seed BH mass distribution. 
% The background UV radiation $$ is calculated by semi-analytical arguments of men
%The merger heating and enhancement of background Lyman-Werner radiation intensity $\jlw$ 
%due to halo clustering is taken into account to shape the gas thermal evolution.

@@

@@

%%%%%%%
%    Fig. 2
%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=150mm]{scheme1.png}
\caption{\blue{
Schematical demonstration of the procedure to calculate the evolution of BH mass distribution from the earliest BH seed to $z=6$.
The checkpoints are spaced evenly with time interval $\tau$.
From time $T_0$ to $T_0+\tau$, 
the BH mass distribution evolves with $\Delta t=\tau$.
For a newly formed seed BH $M_{\rm i}$, the probability distribution function of BH mass is a delta function, 
and the growth time is $\Delta t = T_0+\tau-t_{\rm i}$.
The probability of a BH with mass $M_0$ (or $M_{\rm i}$) to grow to $M$ after $\Delta t$ is drew from the 
ERDF following Eq.~(\ref{eq:Pl}). 
Thus the evolution of BH mass distribution is reshaped iteratively 
plus taking account of the newly formed seeds if there are any.}
}
\label{fig:scheme}
\vspace{5mm}
\end{figure*}

@@

%%%%%%%
%    Fig. 3
%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=180mm]{contour.png} % contour.pdf
\caption{
Two dimensional posterior distribution of all the model parameters with $\fseed=0.1$ and 0.01, 
along with the marginalized one dimensional projection.
% and the two dimensional distribution between every two parameters. 
The distribution of three parameters ($\tau$, $\lambda_0$, and $\alpha$) show single peaks,
displaying the convergence of those parameters.
The distribution of $\log\delta$ is relatively broader but suggests that the value of $\delta$
is sufficiently low and thus the mass-dependent growth model is excluded.
}
\label{fig:contour}
\vspace{5mm}
\end{figure*}
%

@@

% 3. posterior
% Then evaluating the likelihood of the dataset reproducing the observed BHMF and QLF at $z=$ 6, 
% the Markov chain of samples converges to the best-fit (maximum likelihood) parameters.
Following the prior distribution functions, the four parameters are sampled to generate the synthetic 
BHMF and QLF dataset, which are compared to the observational dataset. 
The discrepancy between the model and observational data is evaluated with a $\chi^2$-value defined by
%%%%%%%
%    Eq. 8
%%%%%%%
\begin{align}
  \chi^2 = \sum_{x,i}
  \frac{\left(\log{\Phi^{\rm mod}_{x,i}} - \log{\Phi^{\rm obs}_{x,i}}\right)^2}{(\log{\Phi^{\rm err}_{x,i}})^2},
\end{align}
where $\Phi^{\rm mod}_{x,i}$ and $\Phi^{\rm obs}_{x,i}$ represent the modeled and observational value 
on the \textit{i-}th bin of the BHMF ($x=\Mbh$) and QLF ($x=\Muv$), and $\Phi^{\rm err}_{x,i}$ is the corresponding error.
% on the \textit{i-}th bin of the BHMF ($j=1$) and QLF ($j=2$), and $\Phi^{\rm err}_{i,j}$ is the corresponding error.
Adopting 10 mass bins and 12 magnitude bins, we treat the errors from comparison both in the BHMF and QLF
with a nearly equal weight.
%te that the equal constraints from the BHMF and QLF are realized by extracting
%comparable numbers of data points from $\Phi_\mathrm{M}$ and $\Phi_\mathrm{\Muv}$.
% the constraint is actually insensitive to the choice of bin numbers. 
The posterior distribution of the MCMC fitting is stabilized around the minimum value of $\chi^2$ (highest likelihood), 
with the models best reproducing the $z=$ 6 BHMF and QLF.

@@

%%%%%%%
%    Table. 1
%%%%%%%
\begin{deluxetable}{l|ccccccc}%[t!]
\tablecolumns{1}
\renewcommand\thetable{1} %! fix indexing
\tablewidth{0pt} 
\tablecaption{Best-fit Parameters \label{tab:best_fits}} %!!!!
\tablehead{
  \colhead{$\fseed$}  & \colhead{1} & \colhead{0.1} &  \colhead{0.01}
}
\startdata
% $\fseed$    &      1 &      0.1  &  0.01  \\
% \hline \hline
$\tau$      &  23.13 &   20.07   & 18.76  \\
$\delta$    &  0.001 &   0.001   & 0.055  \\
$\lambda_0$ &  0.96  &   0.89    &  0.87  \\    
$\alpha$    & -0.06  & 0.12   &  0.20  \\
\enddata
% \tablecomments{
% The best-fit parameters optimized by MCMC sampling for the cases of $\fseed=1$, 0.1, and 0.01.
% }
\end{deluxetable}

@@

%%%%%%%
%    Fig. 4
%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=145mm]{fits_MF.png}
\caption{
The BH mass function at $z=6$ with the best-fit parameters (black curve) and the $1\sigma$ spread (shaded region) for the cases with $\fseed=0.1$ (left) and $0.01$ (right). 
For comparison, the BHMF inferred by \citetalias{2010AJ....140..546W} is overlaid (blue curve and shaded region).
The data is used for the model parameter fitting.
%.The blue line shows the BHMF derived by \citet{2010AJ....140..546W}.
% error of observation
}
\label{fig:fitmf}
\end{figure*}

%%%%%%%
%    Fig. 5
%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=145mm]{fits_LF.png}
\caption{
The unobscured quasar luminosity function at $z=6$ with the best-fit parameters (black solid line) and the $1\sigma$ spread for the cases with $\fseed=0.1$ (left) and $0.01$ (right).
The observed data (blue symbols) with error bars are taken from \citetalias{2018ApJ...869..150M} and is used for the model parameter fitting.
The upper bound of faint quasar number density is derived from the cumulative QLF presented in \citet{2022NatAs...6..850J}.
The X-ray QLF data in the faint end inferred from the $z\sim 5-6.1$ quasar sample in \citet{2019ApJ...884...19G} are denoted by orange symbols,
note that we conduct an extrapolation with quasar number densities scaling with $10^{-0.72\Delta z}$ at $z$ in $5-6$ \citep{2016ApJ...833..222J}.
For comparison, we show the intrinsic QLF predicted by our model in dashed lines.
}
\label{fig:fitlf}
\vspace{2mm}
\end{figure*}

@@

Fig.~\ref{fig:fitlf} presents the QLFs at $z=6$ reproduced by the best-fit parameters as well as the $1\sigma$ spreads.
The solid and dashed curve is the {\it unobscured} QLF %(excluding a contribution from obscured quasars)
and the {\it intrinsic} QLF (including both obscured and unobscured quasars), respectively.
For comparison, we overlay the observational data used for the fitting and the QLF parametrized with 
a double power-law shape obtained by \citetalias{2018ApJ...869..150M} (dotted curve).
The best-fit unobscured QLF is consistent with the observed one at $-29~{\rm mag} \lesssim \Muv \lesssim -25~{\rm mag}$,
but overproduces fainter quasars at $\Muv\gtrsim -24~{\rm mag}$.
%, reflecting both the less constraints from observational data in the faint end,
%and the difficulty of reproducing such a small population of low-luminosity quasars based on our seeding and growth model.
As seen in the BHMF, the difference between the model and observational data at the faint end becomes smaller with
the lower seeding fraction.
Since the progenitors of those faint quasars were seed BHs located around the peak of the mass distribution,
those BHs naturally produces a flatter slope of the QLF at the faint end.
In addition, we show upper bounds of the faint quasar abundance at $\Muv\gtrsim -22~{\rm mag}$ \citep[green symbols;][]{2022NatAs...6..850J}
%, an upper limit of the faint end QLF is provided, % by non-detection of faint quasars
based on a search for quasars in the Hubble Space Telescope (HST) imaging fields.
Our unobscured QLF lies within the constraints.

Quasar observations in X-rays also provide 
useful constraints on the intrinsic population because of the different quasar selection criteria and less obscuration effects in X-rays.
Based on the X-ray selected faint quasars, \cite{2015A&A...578A..83G,2019ApJ...884...19G} reported higher number densities of faint quasars
compared to those extrapolated from the rest-UV observations (e.g., the dotted curve by \citetalias{2018ApJ...869..150M}) 
down to the faint end of $\Muv\gtrsim -22~{\rm mag}$ (orange symbols).
%In addition to the systematic difference in the quasar identification methods,
%the X-ray observation should suffer less dust obscuration (see \S~\ref{sec:LF}),
%thus detect more quasars in the faint end.
%We also show the intrinsic QLF predicted by our best-fit parameters,
%which is without imposing obscuration thus representing the upper bound of detectable quasar numbers.
Our best-fitted intrinsic QLF (dashed curve) is broadly consistent with the faint-end QLF from the X-ray observations within the errors,
though the model predicts slightly lower numbers at $\Muv\gtrsim -21$ mag.
% outshining effect
From another point of view, the discrepancy between the UV and X-ray QLFs at the faint end would be explained by the fact that 
only bright quasars outshining their host galaxies can be identified as point-like sources in the rest-UV quasar selection 
(\citetalias{2018ApJ...869..150M}; \citealt{2020MNRAS.495.2135N,2021MNRAS.502.2757O,2022arXiv220712282K}).
However, at the faint end at $\Muv\gtrsim -24$ mag, extended sources such as the brightest galaxies at $z\sim 6$ 
dominate over quasars in number at the same UV magnitude \citep{2022ApJS..259...20H}.
Thus, the current observations might miss quasars embedded in extended galaxy populations. 
% We do not take into account of this effect in the QLF correction,
% but we notice this important caveat and provide the galaxy number density information 
% in the higher-$z$ QLF predictions (see \S~\ref{sec:qlfglf}).
We leave this issue as an important caveat in high-$z$ quasar observations.

@@

%%%%%%%
%    Fig. 6
%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=175mm]{BHMF_rhoz1.png}
\caption{
{\it Left panel}: BH mass functions predicted by the best-fit models at $z=6-10$,
including a fraction of $\fseed$ seeds evolving by our model assumption and the remaining ($1-\fseed$) fraction with no subsequent growth.
The solid (dashed) lines denote the cases with $\fseed=0.1$ (0.01).
{\it Right panel}: the cumulative mass density evolution for BHs in different mass ranges in a comoving volume, 
evaluated by the integration of the BHMF shown in the left panel.
The dashed-dotted line and open square symbol show the results from \citet{2022MNRAS.513..670N}.
The upper bound of accreted BH mass density from \citet{2013ApJ...778..130T} is denoted by red triangles.
% The evolutionary trend can be approximated as $\rho_\bullet (z) \propto 10^{k_M z}$, where $k_M\simeq -0.68$.
}
\label{fig:BHMF_rhoz}
\vspace{0mm}
\end{figure*}

@@

%%%%%%%
%    Table. 2 
%%%%%%%
\begin{deluxetable*}{lccccccc}%[t!]
\tablecolumns{1}
\renewcommand\thetable{2} %! fix indexing
\tablewidth{0pt} 
\tablecaption{Detection Number of Quasars Estimated for Future Surveys}
\tablehead{
  \colhead{Telescope} & \colhead{ Survey area} & \colhead{Photo. depth} &  \colhead{$N(z\sim 6)$} & \colhead{$N(z\sim 7)$}& \colhead{$N(z\sim 8)$}&  \colhead{$N(z\sim 9)$}&  \colhead{$N(z\sim 10)$}\\
  \colhead{} & \colhead{(\rm{deg}$^2$)} & \colhead{(AB mag)} & & & & & 
}
\startdata
Wide RST    &  2000  & $YJH$ 27   & $6759_{-3185}^{+15} $ &  $1514_{-558}^{+169}$ &  $349_{-131}^{53}$ & $85_{-48}^{+23}$ & $25^{+12}_{-15}$  \\
Wide Euclid & 15000  & $YJH$ 24   & $6654^{+95}_{-2282}$  &  $708^{+213}_{-237}$ &   $93^{+66}_{-56}$ & $14^{+32}_{-27}$ & --  \\
\enddata
\label{tab:N_detect}
\tablecomments{
Based on the 5$\sigma$ point source detection limits of the $YJH$ bands in the two surveys,
the number of quasars at each redshift are predicted
with the assumption of the survey volume with $\Delta z =1$ and 100\% completeness.
}
\end{deluxetable*}

@@

\blue{
Moreover, some individual quasar observations give exceptionally small proximity zone sizes,
indicating short lifetime of their luminous phase $t_{\rm Q} \lesssim 10^4$ yr.
The short quasar lifetime may be due to the young age of quasars ionizing the surrounding IGM \citep[e.g.,][]{2020ApJ...903...60I,2021ApJ...917...38E}.
Our episodic quasar growth model agrees with this argument,
besides alleviates the growth tension of SMBHs by episodic accretion bursts.
In other words, the lifetime of luminous quasar phase is the cumulative time of accreting activity in the current episode.
}

>>>>

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
The early evolution of the quasar luminosity function (QLF) and black hole mass function (BHMF) encodes 
key information on the physics determining the radiative and accretion processes of supermassive 
black holes (SMBHs) in high-redshift quasars.
Although the shape of the QLF has been constrained by recent wide and deep surveys,
it still remains challenging to develop a theoretical model that explains its redshift evolution with BH growth self-consistently.
In this study, based on a semi-analytical model for the formation and growth of seed BHs
in quasar progenitor halos,
we construct the QLF and BHMF of the early BH population that experiences multiple accretion bursts,
in each of which a constant Eddington ratio is assigned following a Schechter distribution function.
Our best-fit model to reproduce the observed QLF and BHMF at $z\simeq 6$
suggests that several episodes of super-Eddington accretion occur
and each of them lasts $\tau \simeq 20-30$ Myr.
The average duty cycle in super-Eddington phases of BHs that reach $\gtrsim 10^8~\Msun$ by $z\simeq 6$ is as high as $\simeq 15~\%$,
which is nearly twice that of the entire population.
We also find that the observed Eddington-ratio distribution function is skewed to a log-normal shape from the intrinsic Shechter-like function
owing to detection limits of quasar surveys.
The predicted redshift evolution of both QLF and BHMF of the earliest BHs suggests a rapid decay of their number and mass
toward higher redshifts of $z>6$, compared to galaxy stellar mass.
This will be unveiled by future deep and wide surveys with the James Webb Space Telescope, Roman Space Telescope, and Euclid.
\end{abstract}

@@

We consider seed BH formation in the main progenitors of quasar host galaxies, following a semi-analytical model established by \citet{2021ApJ...917...60L}.
In the highly-biased, overdense regions of the universe, those progenitor halos are likely irradiated by intense H$_2$-photodissociating radiation 
from nearby star-forming galaxies and heat the interior gas by successive mergers. 
The two effects counteracting H$_2$ formation and cooling prevent gas collapse and delay prior star formation \citep[e.g.,][]{2014MNRAS.445..107V,2019Natur.566...85W}.
Under such peculiar circumstances, massive clouds collapse to the halo centers at high accretion rates and leave massive stars (presumably seed BHs) behind.
Our model takes into account various key physical processes on star formation such as halo merger heating, radiative cooling, and
(photo-)chemical reaction networks.
%, and gravitational instability of the central gaseous cores.
The time evolution of the gas dynamics and thermal states is calculated in a self-consistent way with the halo assembly history.
%the gas dynamical and thermal evolution is regulated by the parent halo growth histories, and the environmental streaming velocities.
%
%Next we introduce our seeding model where the gas dynamical and thermal evolution is regulated by the parent halo growth histories, 
%and the environmental streaming velocities.
%In \citet{2021ApJ...917...60L}, we propose an semi-analytical model which 
%traces the gas evolution along its parent halo growth and yields the seed BH mass distribution. 
% The background UV radiation $$ is calculated by semi-analytical arguments of men
%The merger heating and enhancement of background Lyman-Werner radiation intensity $\jlw$ 
%due to halo clustering is taken into account to shape the gas thermal evolution.

@@

\orange{
Lastly, we note that in each parent halo, we consider purely the PopIII stellar remnants in the seed BH mass function construction,
neglecting the BHs formed by the second-generation PopII stars.
Formed in metal-enriched environments, PopII stellar remnants are more abundant but typically less massive.
They are likely to dominate the low mass end of the BHMF at $z\sim 6$,
thus not significantly contribute to the observable QLF and BHMF \citep{2022MNRAS.511..616T}.
}

@@

%%%%%%%
%    Fig. 2
%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=150mm]{scheme1.pdf}
\caption{\orange{
Schematical demonstration of the procedure to calculate the evolution of BH mass distribution from the earliest BH seed to $z=6$.
The checkpoints are spaced evenly with time interval $\tau$.
From time $T_0$ to $T_0+\tau$, 
the BH mass distribution evolves with $\Delta t=\tau$.
For a newly formed seed BH $M_{\rm i}$, the probability distribution function of BH mass is a delta function, 
and the growth time is $\Delta t = T_0+\tau-t_{\rm i}$.
The probability of a BH with mass $M_0$ (or $M_{\rm i}$) to grow to $M$ after $\Delta t$ is drew from the 
ERDF following Eq.~(\ref{eq:Pl}). 
Thus the evolution of BH mass distribution is reshaped iteratively 
plus taking account of the newly formed seeds if there are any.}
}
\label{fig:scheme}
\vspace{5mm}
\end{figure*}

@@

%%%%%%%
%    Fig. 3
%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=85mm]{f1_corner.pdf}
\includegraphics[width=85mm]{f2_corner.pdf}
% \includegraphics[width=180mm]{contour.png}
\caption{
Two dimensional posterior distribution of all the model parameters with $\fseed=0.1$ and 0.01, 
along with the marginalized one dimensional projection.
% and the two dimensional distribution between every two parameters. 
The distribution of three parameters ($\tau$, $\lambda_0$, and $\alpha$) show single peaks,
displaying the convergence of those parameters.
The distribution of $\log\delta$ is relatively broader but suggests that the value of $\delta$
is sufficiently low and thus the mass-dependent growth model is excluded.
}
\label{fig:contour}
\vspace{5mm}
\end{figure*}
%

@@

% 3. posterior
% Then evaluating the likelihood of the dataset reproducing the observed BHMF and QLF at $z=$ 6, 
% the Markov chain of samples converges to the best-fit (maximum likelihood) parameters.
Following the prior distribution functions, the four parameters are sampled to generate the synthetic 
BHMF and QLF dataset, which are compared to the observational dataset. 
The discrepancy between the model and observational data is evaluated with a $\chi^2$-value defined by
%%%%%%%
%    Eq. 8
%%%%%%%
\begin{align}
  \label{eq:chi2}
  \chi^2 = \sum_{x,i}
  \frac{\left(\log{\Phi^{\rm mod}_{x,i}} - \log{\Phi^{\rm obs}_{x,i}}\right)^2}{(\log{\Phi^{\rm err}_{x,i}})^2},
\end{align}
where $\Phi^{\rm mod}_{x,i}$ and $\Phi^{\rm obs}_{x,i}$ represent the modeled and observational value 
on the \textit{i-}th bin of the BHMF ($x=\Mbh$) and QLF ($x=\Muv$), and $\Phi^{\rm err}_{x,i}$ is the corresponding error.
% on the \textit{i-}th bin of the BHMF ($j=1$) and QLF ($j=2$), and $\Phi^{\rm err}_{i,j}$ is the corresponding error.
Adopting 10 mass bins and 12 magnitude bins, we treat the errors from comparison both in the BHMF and QLF
with a nearly equal weight.
%te that the equal constraints from the BHMF and QLF are realized by extracting
%comparable numbers of data points from $\Phi_\mathrm{M}$ and $\Phi_\mathrm{\Muv}$.
% the constraint is actually insensitive to the choice of bin numbers. 
The posterior distribution of the MCMC fitting is stabilized around the minimum value of $\chi^2$ (highest likelihood), 
with the models best reproducing the $z=$ 6 BHMF and QLF.

@@

%%%%%%%
%    Table. 1
%%%%%%%
\begin{deluxetable}{lcccccc}%[t!]
\tablecolumns{1}
\renewcommand\thetable{1} %! fix indexing
\tablewidth{0pt} 
\tablecaption{Best-fit Parameters \label{tab:best_fits}} %!!!!
\tablehead{
  \colhead{$\fseed$}  & \colhead{$\tau$ (Myr)} & \colhead{$\delta$} &  \colhead{$\lambda_0$}  & \colhead{$\alpha$}  & \colhead{$\chi^2$}
}
\startdata
1.0    &  23.13 &    0.001    &   0.96    &   -0.06    &    15.77   \\
0.1    &  20.07 &    0.001    &   0.89    &    0.12    &    8.27    \\
0.01   &  18.76 &    0.055    &   0.87    &    0.20    &    6.21    \\
\enddata
\tablecomments{
The best-fit parameters optimized by MCMC sampling for different cases of $\fseed$.
The fitting performance is characterized by the relative magnitudes of the $\chi^2$ value.% (see Eq.~\ref{eq:chi2}).
}
\end{deluxetable}

@@

%%%%%%%
%    Fig. 4
%%%%%%%
\begin{figure*}
\centering
% \includegraphics[width=145mm]{fits_MF.png}
\includegraphics[width=85mm]{f1ndraw60MF_spread.pdf}
\includegraphics[width=85mm]{f2ndraw60MF_spread.pdf}
\caption{
The BH mass function at $z=6$ with the best-fit parameters (black curve) and the $1\sigma$ spread (shaded region) for the cases with $\fseed=0.1$ (left) and $0.01$ (right). 
For comparison, the BHMF inferred by \citetalias{2010AJ....140..546W} is overlaid (blue curve and shaded region).
The data is used for the model parameter fitting.
%.The blue line shows the BHMF derived by \citet{2010AJ....140..546W}.
% error of observation
}
\label{fig:fitmf}
\end{figure*}

%%%%%%%
%    Fig. 5
%%%%%%%
\begin{figure*}
\centering
% \includegraphics[width=145mm]{fits_LF.png}
\includegraphics[width=85mm]{f1ndraw60LF_spread.pdf}
\includegraphics[width=85mm]{f2ndraw60LF_spread.pdf}
\caption{
The unobscured quasar luminosity function at $z=6$ with the best-fit parameters (black solid line) and the $1\sigma$ spread for the cases with $\fseed=0.1$ (left) and $0.01$ (right).
The observed data (blue symbols) with error bars are taken from \citetalias{2018ApJ...869..150M} and is used for the model parameter fitting.
The upper bound of faint quasar number density is derived from the cumulative QLF presented in \citet{2022NatAs...6..850J}.
The X-ray QLF data in the faint end inferred from the $z\sim 5-6.1$ quasar sample in \citet{2019ApJ...884...19G} are denoted by orange symbols,
note that we conduct an extrapolation with quasar number densities scaling with $10^{-0.72\Delta z}$ at $z$ in $5-6$ \citep{2016ApJ...833..222J}.
For comparison, we show the intrinsic QLF predicted by our model in dashed lines.
}
\label{fig:fitlf}
\vspace{2mm}
\end{figure*}

@@

Fig.~\ref{fig:fitlf} presents the QLFs at $z=6$ reproduced by the best-fit parameters as well as the $1\sigma$ spreads.
The solid and dashed curve is the {\it unobscured} QLF %(excluding a contribution from obscured quasars)
and the {\it intrinsic} QLF (including both obscured and unobscured quasars), respectively.
For comparison, we overlay the observational data used for the fitting and the QLF parametrized with 
a double power-law shape obtained by \citetalias{2018ApJ...869..150M} (dotted curve).
The best-fit unobscured QLF is consistent with the observed one at $-29~{\rm mag} \lesssim \Muv \lesssim -25~{\rm mag}$,
but overproduces fainter quasars at $\Muv\gtrsim -24~{\rm mag}$.
%, reflecting both the less constraints from observational data in the faint end,
%and the difficulty of reproducing such a small population of low-luminosity quasars based on our seeding and growth model.
As seen in the BHMF, the difference between the model and observational data at the faint end becomes smaller with
the lower seeding fraction.
Since the progenitors of those faint quasars were seed BHs located around the peak of the mass distribution,
those BHs naturally produces a flatter slope of the QLF at the faint end.
In addition, we show upper bounds of the faint quasar abundance at $\Muv\gtrsim -22~{\rm mag}$ \citep[green symbols;][]{2022NatAs...6..850J}
%, an upper limit of the faint end QLF is provided, % by non-detection of faint quasars
based on a search for point sources in the survey fields of the Hubble Space Telescope (HST).
Our unobscured QLF lies within the constraints.

Quasar observations in X-rays also provide 
useful constraints on the intrinsic population because of the different quasar selection criteria and less obscuration effects in X-rays.
Based on the X-ray selected faint quasars, \cite{2015A&A...578A..83G,2019ApJ...884...19G} reported higher number densities of faint quasars
compared to those extrapolated from the rest-UV observations (e.g., the dotted curve by \citetalias{2018ApJ...869..150M}) 
down to the faint end of $\Muv\gtrsim -22~{\rm mag}$ (orange symbols).
%In addition to the systematic difference in the quasar identification methods,
%the X-ray observation should suffer less dust obscuration (see \S~\ref{sec:LF}),
%thus detect more quasars in the faint end.
%We also show the intrinsic QLF predicted by our best-fit parameters,
%which is without imposing obscuration thus representing the upper bound of detectable quasar numbers.
Our best-fitted intrinsic QLF (dashed curve) is broadly consistent with the faint-end QLF from the X-ray observations within the errors,
though the model predicts slightly lower numbers at $\Muv\gtrsim -21$ mag.
% outshining effect
From another point of view, the discrepancy between the UV and X-ray QLFs at the faint end would be explained by the fact that 
only bright quasars outshining their host galaxies can be identified as point-like sources in the rest-UV quasar selection 
(\citetalias{2018ApJ...869..150M}; \citealt{2020MNRAS.495.2135N,2020MNRAS.494.1771A,2021MNRAS.502.2757O,2021MNRAS.502..662B,2022arXiv220712282K}).
However, at the faint end at $\Muv\gtrsim -24$ mag, extended sources such as the brightest galaxies at $z\sim 6$ 
dominate over quasars in number at the same UV magnitude \citep{2022ApJS..259...20H}.
Thus, the current observations might miss quasars embedded in extended galaxy populations. 
% We do not take into account of this effect in the QLF correction,
% but we notice this important caveat and provide the galaxy number density information 
% in the higher-$z$ QLF predictions (see \S~\ref{sec:qlfglf}).
We leave this issue as an important caveat in high-$z$ quasar observations.

@@

%%%%%%%
%    Fig. 6
%%%%%%%
\begin{figure*}
\centering
% \includegraphics[width=175mm]{BHMF_rhoz1.png}
\includegraphics[width=85mm]{MF.pdf}
\includegraphics[width=85mm]{rhoM_comb.pdf}
\caption{
{\it Left panel}: BH mass functions predicted by the best-fit models at $z=6-10$,
including a fraction of $\fseed$ seeds evolving by our model assumption and the remaining ($1-\fseed$) fraction with no subsequent growth.
The solid (dashed) lines denote the cases with $\fseed=0.1$ (0.01).
{\it Right panel}: the cumulative mass density evolution for BHs in different mass ranges in a comoving volume, 
evaluated by the integration of the BHMF shown in the left panel.
The dashed-dotted line and open square symbols show the results from \citet{2022MNRAS.513..670N}.
The upper bound of accreted BH mass density from \citet{2013ApJ...778..130T} is denoted by black circles.
% The evolutionary trend can be approximated as $\rho_\bullet (z) \propto 10^{k_M z}$, where $k_M\simeq -0.68$.
}
\label{fig:BHMF_rhoz}
\vspace{0mm}
\end{figure*}

@@

%%%%%%%
%    Table. 2 
%%%%%%%
\begin{deluxetable*}{lccccccc}%[t!]
\tablecolumns{1}
\renewcommand\thetable{2} %! fix indexing
\tablewidth{0pt} 
\tablecaption{Detection Number of Quasars Estimated for Future Surveys \label{tab:N_detect}} %! reference indexing
\tablehead{
  \colhead{Telescope} & \colhead{ Survey area} & \colhead{Photo. depth} &  \colhead{$N(z\sim 6)$} & \colhead{$N(z\sim 7)$}& \colhead{$N(z\sim 8)$}&  \colhead{$N(z\sim 9)$}&  \colhead{$N(z\sim 10)$}\\
  \colhead{} & \colhead{(\rm{deg}$^2$)} & \colhead{(AB mag)} & & & & & 
}
\startdata
Wide RST    &  2000  & $YJH$  27   & $6759_{-3185}^{+15} $ &  $1514_{-558}^{+169}$ &  $349_{-131}^{53}$ & $85_{-48}^{+23}$ & $25^{+12}_{-15}$  \\
Wide RST    &  2000  & $\Muv=-26$ & $21_{-5}^{+13} $      &  $2_{-7}^{+8}$        &  --  & -- & --  \\
Wide Euclid & 15000  & $YJH$  24   & $6654^{+95}_{-2282}$  &  $708^{+213}_{-237}$ &   $93^{+66}_{-56}$ & $14^{+32}_{-27}$ & --  \\
Wide Euclid & 15000  & $YJH$  23   & $2480^{+188}_{-614}$  &  $207^{+144}_{-114}$ &   $23^{+27}_{-19}$ & $3^{+12}_{-10}$ & --  \\
Wide Euclid & 15000  & $\Muv=-26$ & $154^{+100}_{-38}$  &  $12^{+59}_{-50}$ &   $1^{+4}_{-3}$ & -- & --  \\
\enddata
\tablecomments{
Predicted number of quasars at each redshift in the two surveys, with the assumption of the survey volume with $\Delta z =1$ and 100\% completeness.
The integration of QLF starts from the 5$\sigma$ point source detection limits of the $YJH$ bands in each survey, 
and the bright magnitude with $\Muv=-26$ mag.
In addition, detection limit with 22 mag in the $YJH$ bands is examined for Euclid,
mimicking the effect of quasar selection function.% following \citet{2019A&A...631A..85E}.
}
\end{deluxetable*}

@@

\blue{
Moreover, some individual quasar observations give exceptionally small proximity zone sizes,
indicating short lifetime of their luminous phase $t_{\rm Q} \lesssim 10^4$ yr.
The short quasar lifetime may be due to the young age of quasars ionizing the surrounding IGM \citep[e.g.,][]{2017ApJ...840...24E,2020ApJ...903...60I,2021ApJ...917...38E}.
Our episodic quasar growth model agrees with this argument,
besides alleviates the growth tension of SMBHs by episodic accretion bursts.
In other words, the lifetime of luminous quasar phase is the cumulative time of accreting activity in the current episode.
}
